Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics

### 2019-08-05 review/response.Rmd

### Associate Editor:

The submission has been evaluated by two reviewers. The manuscript is well written and organized. I have some more philosophical issues in addition to those raised by the two reviewers.

The authors need to convince readers (users) this kind of environment is good for daily EDA for the real world of big data/data science.

**We have added a paragraph to the introduction to explain the approach for higher dimensional problems, as follows:**

**PP can also be used to explore data with many more variables, data that is commonly seen today. For example, [@Cook:2018mvr] explores a 56-dimensional parameter space, by first reducing the number of dimensions to the first six principal components, before applying projection pursuit. There is not a single solution for working with high-dimensional data. PCA was appropriate for this problem because reducing to principal component space removed the linear dependencies but preserved the nonlinear relationships that were interesting to discover. Some projection pursuit indexes incorporation penalty terms to naturally deal with noise dimensions. Having an efficient optimizer can also important with increasing dimensionality, because there are many more dimensions to search.**

@@

Example(s) with very large number (at least hundreds) of dimensions should be included in the study.

**The simulation study and the applications reveal that the current methods are efficient only in smaller parameter spaces. There are several components that would be required to scale it to examples with hundreds of dimensions: (1) fast computational engine, (2) an efficient optimiser, (3) new PP indexes that effectively use penalty terms to automatically elliminate nuisance dimensions. The first two would require major implementation work. The methods explained in this paper would be useful for assessing and evaluating new PP indexes. The recommended approach to high-dimensions is to first reduce the dimensionality using PCA, before applying PP. We have explained a sentence to this effect in the discussion section.**

@@

Who are the target users of this environment? Do you recommend naive users (users without much EDA experience) to get hand on the proposed system?

How do you encourage users to try your system?

**The target audience consists of two groups. First we hope that researchers developing new index functions, who will find our methods useful to diagnose their new indexes. We have implemented out methods in the R package, spinebil, to make this an attractive prospect to other researchers. The second group would be people interested in applying the indexes to explore nonlinear structure in their data. This can be done directly using the guided tour, in the tourr package, using code similarly to that distributed with this paper. We also encourage novice R users to try this through the tour GUI available in the galahr package. Both packages are now referenced in the discussion.**

### 2019-08-05 scagnostics.Rmd
working on response to reviewers

\textcolor{red}{The same methods can also be used to explore larger datasets with hundreds of dimensions, as common in real world examples. In these scenarios it is best to reduce the number of dimensions before applying projection pursuit. For example in [@Cook:2018mvr] a 56 dimensional parameter space was reduced to the first six principal components before applying tour methods for the identification of multivariate relationships, revealing previously overlooked features. On the other hand, applying projection pursuit directly to the high-dimensional dataset is generally limited by the availability of an efficient optimiser.
}

@@

```{r load}
library(tourr)
library(tidyverse)
library(scagnostics)
library(gridExtra)
library(tictoc) #timer
library(mbgraphic) #Katrins package
library(GGally)
library(geozoo)
library(minerva) #MINE indices
library(kableExtra)
library(forcats)
```

@@

All of the indexes fall short against some aspect of the ideal properties of smoothness, squintability, flexibility, rotation invariance and speed. The paper describes how these properties can be assessed using tour methodology. Some potential fixes for the indexes are discussed but there is scope for further developing the new indexes.
\textcolor{red}{We recommend to use the spinebil [XXX REF] package when developing new indexes. It includes the functionalities needed to reproduce the assessements presented in this work. While the current focus is on two-dimensional index functions, the methodology introduced here directly applies to the assessment of index functions where $d>2$.}

The work also reveals inadequacies in the tour optimization algorithm, that may benefit from newly developed techniques and software tools. Exploring this area would help improve the guided tours.
\textcolor{red}{In particular these developments would enable the application of guided tours to datasets with large number of dimensions.}

@@

>>>>

\textcolor{red}{PP can also be used to explore data with many more variables, data that is commonly seen today. For example, [@Cook:2018mvr] explores a 56-dimensional parameter space, by first reducing the number of dimensions to the first six principal components, before applying projection pursuit. There is not a single solution for working with high-dimensional data. PCA was appropriate for this problem because reducing to principal component space removed the linear dependencies but preserved the nonlinear relationships that were interesting to discover. Some projection pursuit indexes incorporation penalty terms to naturally deal with noise dimensions. Having an efficient optimizer can also important with increasing dimensionality, because there are many more dimensions to search.}

@@

```{r load}
library(tourr)
library(tidyverse)
library(binostics)
library(gridExtra)
library(tictoc) #timer
library(mbgraphic) #Katrins package
library(GGally)
library(geozoo)
library(minerva) #MINE indices
library(kableExtra)
library(forcats)
```

@@

All of the indexes fall short against some aspect of the ideal properties of smoothness, squintability, flexibility, rotation invariance and speed. The paper describes how these properties can be assessed using tour methodology. Some potential fixes for the indexes are discussed but there is scope for further developing the new indexes.
\textcolor{red}{We recommend to use the spinebil [XXX REF] package when developing new indexes. It includes the functionalities needed to reproduce the assessements presented in this paper. While the current focus is on two-dimensional index functions, indexes in the tourr package apply to arbitrary projection dimension, and the methodology introduced here could be applied to the assessment of index functions where $d>2$.}

The work also reveals inadequacies in the tour optimization algorithm, that may benefit from newly developed techniques and software tools. Exploring this area would help improve the guided tours.
\textcolor{red}{As new optimization techniques become available, adapting these to the guided tour would extend the technique to a broader range of problems. The current recommended approach is to first reduce the dimensionality, for example by using PCA, taking care to preserve nonlinear structure, prior to applying PP.}

@@

# Supplementary material

- R code for paper
- spinebil
- galahr
- outline of appendix
